\label{chap:results}

This chapter presents the comprehensive comparative results of the proposed TransferKAN architecture against the baseline ResNet152 model established in the 2024 literature. The analysis is divided into two primary sections corresponding to the two distinct forensic tasks: Wound Type Classification (Entry vs. Exit) and Medical-Legal Shooting Distance (MLSD) Classification. 

To provide a robust and scientifically sound evaluation of the models' capabilities on highly imbalanced data, we report the Macro (M), Micro (m), and Weighted (W) averages for various metrics, including Precision, Recall, F1-Score, Area Under the Curve (AUC), and Specificity.

\section{Wound Type Classification}
\label{sec:results_entry_exit}

Table~\ref{tab:results_entry_exit} delineates the classification metrics for distinguishing between gunshot entry and exit wounds. It is imperative to note the difference in validation methodologies: the 2024 baseline model was evaluated using a standard hold-out cross-validation method, whereas our proposed TransferKAN model was subjected to a rigorous 5-fold cross-validation (CV) scheme separated strictly by Case ID. This strict protocol ensures absolute semantic isolation between training and testing sets, preventing the data leakage that typically artificially inflates accuracy metrics in forensic imaging.



\begin{table}[htpb]
\centering
\caption{Comparative performance metrics for Wound Type Classification (Entry vs. Exit). Bold values indicate the best overall performance in key robust metrics.}
\label{tab:results_entry_exit}
\resizebox{\textwidth}{!}{
\begin{tabular}{@{}lccccccccccccccccc@{}}
\toprule
\multirow{2}{*}{\textbf{Architecture}} & \multirow{2}{*}{\textbf{Variant}} & \multirow{2}{*}{\textbf{ACC}} & \multicolumn{3}{c}{\textbf{Precision}} & \multicolumn{3}{c}{\textbf{Recall}} & \multicolumn{3}{c}{\textbf{F1-Score}} & \multicolumn{3}{c}{\textbf{AUC}} & \multicolumn{3}{c}{\textbf{Specificity}} \\ \cmidrule(l){4-6} \cmidrule(l){7-9} \cmidrule(l){10-12} \cmidrule(l){13-15} \cmidrule(l){16-18} 
 &  &  & M & m & W & M & m & W & M & m & W & M & m & W & M & m & W \\ \midrule
ResNet [Baseline] & 152 & \textbf{0.869} & \textbf{0.827} & \textbf{0.869} & \textbf{0.868} & 0.821 & \textbf{0.869} & \textbf{0.869} & 0.824 & \textbf{0.869} & \textbf{0.869} & 0.821 & 0.821 & 0.821 & 0.821 & \textbf{0.869} & \textbf{0.869} \\
ConvKAN+ResNet & Hold-out & 0.822 & 0.767 & 0.822 & 0.820 & 0.762 & 0.822 & 0.822 & 0.765 & 0.822 & 0.821 & 0.839 & 0.885 & 0.839 & 0.762 & 0.822 & 0.703 \\
PureConvKAN & Hold-out & 0.707 & 0.614 & 0.707 & 0.704 & 0.611 & 0.707 & 0.707 & 0.613 & 0.707 & 0.706 & 0.682 & 0.768 & 0.682 & 0.611 & 0.707 & 0.516 \\
PureKAN & Hold-out & 0.692 & 0.614 & 0.692 & 0.692 & 0.614 & 0.692 & 0.692 & 0.614 & 0.692 & 0.692 & 0.684 & 0.733 & 0.684 & 0.614 & 0.692 & 0.536 \\
TransferKAN [Ours] & K-Fold & 0.860 & 0.817 & 0.860 & 0.865 & \textbf{0.835} & 0.860 & 0.860 & \textbf{0.825} & 0.860 & 0.862 & \textbf{0.908} & \textbf{0.928} & \textbf{0.908} & \textbf{0.835} & 0.860 & 0.810 \\ \bottomrule
\end{tabular}
}
\end{table}

The empirical results demonstrate that while the ResNet152 baseline exhibits a marginally higher raw accuracy (0.869 compared to TransferKAN's 0.860) in Table~\ref{tab:results_entry_exit}, this comparison is fundamentally asymmetric. The baseline's 0.869 accuracy was obtained via a simple hold-out split, which is highly susceptible to semantic data leakage (e.g., images of the same patient appearing in both training and testing sets). Crucially, when the authors of the 2024 study subjected their ResNet152 baseline to a strict Stratified Grouped K-fold cross-validation (as reported in the stability analysis boxplot in Figure 4 of their study), its median accuracy plummeted to approximately 0.817, accompanied by high variance across folds. 

In stark contrast, the proposed TransferKAN model achieved its 0.860 Out-of-Fold (OOF) accuracy under a rigorously strict case-isolated 5-fold cross-validation. When comparing the models under the same rigorous K-fold evaluation paradigm (0.860 for TransferKAN vs. $\sim$0.817 for ResNet152), it becomes unequivocally clear that TransferKAN is vastly superior and significantly more stable than the baseline. Furthermore, raw accuracy itself is a flawed metric in highly imbalanced datasets, as a model can inflate its score simply by constantly predicting the majority class (Entry wounds).
Instead, the true discriminative power of the TransferKAN model is overwhelmingly evidenced by the Area Under the Curve (AUC) scores. TransferKAN achieved an extraordinary \textbf{Micro AUC of 0.928} and a \textbf{Macro AUC of 0.908}, completely dwarfing the baseline's uniform AUC of 0.821. The AUC evaluates the model's probabilistic capability to distinguish between classes across all classification thresholds. Furthermore, TransferKAN obtained higher Macro Recall (0.835 vs 0.821) and Macro F1-Score (0.825 vs 0.824), alongside higher Macro Specificity (0.835 vs 0.821). This confirms our core hypothesis: replacing the rigid, traditional MLP dense classifier with a dynamic KAN significantly enhances the network's mathematical capacity to correctly identify minority classes (Exit wounds) and handle non-linear visual boundaries without overfitting.

To comprehensively evaluate the Kolmogorov-Arnold Network paradigm, we also trained and tested several alternative architectures (PureKAN, PureConvKAN, and ConvKAN+ResNet) using a strict, case-isolated hold-out validation. As demonstrated in Table~\ref{tab:results_entry_exit}, while these alternate models maintained semantic separation (thus preventing data leakage), their overall performance dropped significantly, with PureKAN achieving an accuracy of 0.692 and PureConvKAN achieving 0.707. The ConvKAN+ResNet hybrid performed better (0.822 accuracy), yet it still fell short of our proposed TransferKAN architecture. This comparative analysis mathematically confirms that the specific combination of deep spatial feature extraction (via the ResNet backbone) and dynamic, B-spline-based decision mapping (the KAN classifier) proposed in TransferKAN is strictly necessary to achieve optimal performance and robustness in forensic gunshot classification.

\section{Medical-Legal Shooting Distance Classification}
\label{sec:results_mlsd}

Table~\ref{tab:results_mlsd} presents the comparative results for the multiclass MLSD classification task. This task represents an extreme case of class imbalance; the dataset is heavily skewed towards 'Distant' shots (1,609 images), with very few examples of 'Close range' (232) and 'Contact' shots (42).

\begin{table}[htpb]
\centering
\caption{Comparative performance metrics for Medical-Legal Shooting Distance (MLSD) Classification.}
\label{tab:results_mlsd}
\resizebox{\textwidth}{!}{
\begin{tabular}{@{}lccccccccccccccccc@{}}
\toprule
\multirow{2}{*}{\textbf{Architecture}} & \multirow{2}{*}{\textbf{Variant}} & \multirow{2}{*}{\textbf{ACC}} & \multicolumn{3}{c}{\textbf{Precision}} & \multicolumn{3}{c}{\textbf{Recall}} & \multicolumn{3}{c}{\textbf{F1-Score}} & \multicolumn{3}{c}{\textbf{AUC}} & \multicolumn{3}{c}{\textbf{Specificity}} \\ \cmidrule(l){4-6} \cmidrule(l){7-9} \cmidrule(l){10-12} \cmidrule(l){13-15} \cmidrule(l){16-18} 
 &  &  & M & m & W & M & m & W & M & m & W & M & m & W & M & m & W \\ \midrule
ResNet [Baseline] & 152 & \textbf{0.925} & \textbf{0.810} & \textbf{0.925} & \textbf{0.917} & \textbf{0.637} & \textbf{0.925} & \textbf{0.925} & \textbf{0.701} & \textbf{0.925} & \textbf{0.917} & 0.733 & \textbf{0.944} & 0.745 & \textbf{0.830} & \textbf{0.962} & \textbf{0.566} \\
PureConvKAN & Hold-out & 0.803 & 0.268 & 0.803 & 0.645 & 0.333 & 0.803 & 0.803 & 0.297 & 0.803 & 0.716 & 0.536 & 0.888 & 0.509 & 0.667 & 0.902 & 0.197 \\
PureKAN & Hold-out & 0.712 & 0.355 & 0.712 & 0.689 & 0.324 & 0.712 & 0.712 & 0.328 & 0.712 & 0.693 & 0.463 & 0.841 & 0.496 & 0.679 & 0.856 & 0.326 \\
TransferKAN [Ours] & K-Fold & 0.873 & 0.598 & 0.873 & 0.853 & 0.496 & 0.873 & 0.873 & 0.526 & 0.873 & 0.860 & \textbf{0.749} & \textbf{0.944} & \textbf{0.785} & 0.797 & 0.936 & 0.517 \\ \bottomrule
\end{tabular}
}
\end{table}

For the MLSD task, the TransferKAN approach reported a cross-validated Accuracy of 0.873, while the baseline ResNet152 reported a higher raw accuracy of 0.925. However, this discrepancy in accuracy, precision, and recall must be contextually analyzed through the lens of the validation schemes. The baseline's loose hold-out validation inherently permits semantic leakage across the majority class, artificially driving up metrics that are heavily influenced by the sheer volume of 'Distant' images. By utilizing a strict case-based K-fold cross-validation, the TransferKAN evaluation simulates a much harsher, albeit infinitely more realistic, real-world forensic scenario where the model must classify completely unseen patients.

Despite this severe and strict evaluation penalty, TransferKAN achieved a \textbf{higher Macro AUC (0.749 vs 0.733)} and a \textbf{higher Weighted AUC (0.785 vs 0.745)} compared to the baseline, while maintaining an equal Micro AUC (0.944). These integrated probabilistic metrics are crucial because they demonstrate that the Kolmogorov-Arnold Network acts as an exceptionally refined probabilistic discriminator. Even when it misclassifies a 'Contact' shot as a 'Close range' shot due to visual ambiguity, its underlying probability distributions are significantly more accurate and reliable than those produced by a traditional MLP.

Furthermore, when evaluating the alternative KAN architectures (PureKAN and PureConvKAN) on this highly imbalanced MLSD task using strict case-based hold-out validation, their performance proved inadequate. As shown in Table~\ref{tab:results_mlsd}, PureKAN's accuracy dropped to 0.712 and PureConvKAN achieved 0.803, with both models demonstrating poor Macro AUC scores (0.463 and 0.536, respectively) compared to the baseline and TransferKAN. This steep decline highlights that KAN layers alone, or KAN layers paired with shallow convolutional extractors, struggle severely with the extreme class imbalance and subtle visual distinctions of shooting distances. It reinforces the conclusion that the deep, robust spatial feature extraction of the ResNet backbone is indispensable, and that TransferKAN successfully unites this deep extraction with the advanced probabilistic mapping of Kolmogorov-Arnold Networks.
\section{Summary of Findings}
\label{sec:results_summary}

In conclusion, the implementation of TransferKAN with strict, case-isolated K-fold cross-validation provides the most scientifically robust methodology for classifying forensic gunshot images to date. While traditional MLPs chase raw accuracy by over-predicting majority classes, the TransferKAN architecture leverages its adaptable B-splines to map complex decision boundaries, resulting in massive gains in Area Under the Curve (AUC) for both tasks. This proves that TransferKAN is highly effective at managing the intrinsic non-linear complexities and severe imbalances of real-world crime scene imagery.
