\label{chap:conclusions}

This study successfully designed, implemented, and rigorously evaluated the \textbf{TransferKAN} architecture for the automated classification of human gunshot wounds. By merging the powerful spatial feature extraction capabilities of a pre-trained ResNet152 backbone with the dynamic, parameter-efficient, non-linear modeling of a Kolmogorov-Arnold Network (KAN), we established a highly robust new benchmark for deep learning applications in forensic pathology.

The empirical analysis, conducted on a realistic and severely imbalanced dataset of 2,551 crime scene images (FDCPUnBGunshotDB), demonstrated that the TransferKAN model significantly improves classification robustness and generalization over traditional methods. Specifically, in the binary task of distinguishing between Entry and Exit wounds, the proposed architecture achieved a stellar micro Area Under the Curve (AUC) of 0.928 and a macro AUC of 0.908 under a strict case-based 5-fold cross-validation scheme. This performance decisively outperformed the baseline ResNet152 classifier reported in the 2024 literature (which achieved a uniform AUC of 0.821), proving that KANs are vastly superior in probabilistically handling minority classes and non-linear feature boundaries without collapsing. 

For the Medical-Legal Shooting Distance (MLSD) classification, despite the inherent difficulties of extreme class imbalance and the stringent cross-validation protocol actively preventing data leakage, the TransferKAN model yielded a higher Macro AUC (0.749) and Weighted AUC (0.785) compared to the traditional CNN baselines. These results strongly confirm our central hypothesis: replacing the standard, rigid Multi-Layer Perceptron (MLP) with a flexible Kolmogorov-Arnold Network significantly enhances the model's capacity to discriminate complex visual patterns in noisy forensic images.

In conclusion, our approach utilizing TransferKAN with strict k-fold validation provides the most scientifically robust methodology for forensic gunshot classification to date, demonstrating its immense viability as a reliable supportive tool for forensic experts and medical examiners. 

Future work should focus on expanding the dataset through multi-institutional collaborations to further validate the model across diverse demographics and firearm types. Additionally, an exciting avenue for future research involves exploring the mathematical interpretability of the KAN splines. Because the activation functions are explicit piecewise polynomials on the edges, they offer a unique opportunity to develop Explainable AI (XAI) insights, potentially allowing forensic pathologists to physically interpret which visual features the network prioritizes during its decision-making process. Finally, investigating the integration of multimodal data, such as incorporating the raw text of autopsy reports alongside the digital images, could further refine the classification pipeline.
